---
title: "Mortage Value in Cyprus"
author: "Lefkios Paikousis"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
  pdf_document:
    latex_engine: xelatex
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(knitr)
library(palmerpenguins)

library(tidyverse)

Sys.setlocale(locale = "greek")

```

```{r read_data, include=TRUE}

path <- "https://www.data.gov.cy/node/3886/download"

dta <- read_csv(path, col_types = c("c", "f", "n", "n"))

# MONTH is coming in as TEXT. We need it as `Date` format

dta$date <- lubridate::my(dta$MONTH)

mortage_year <- unique(lubridate::year(dta$date))

max_tbl <- dta %>% filter(`Total Amount` == max(`Total Amount`))

```

### Data - Year `r mortage_year`

These are data for mortage values for year `r mortage_year` in Cyprus.

The raw data can be downloaded directly from [here](%60r%20path%60)

You can also see the rest of the datasets at the [data.gov.cy](https://www.data.gov.cy/dataset/%CF%83%CF%84%CE%B1%CF%84%CE%B9%CF%83%CF%84%CE%B9%CE%BA%CE%AC-%CF%85%CF%80%CE%BF%CE%B8%CE%B7%CE%BA%CF%8E%CE%BD-%CE%B1%CE%BA%CE%B9%CE%BD%CE%AE%CF%84%CF%89%CE%BD) website


### Explore the raw data

```{r}
DT::datatable(dta)
```


### Descriptives

The biggest total mortage value **(`r format(max_tbl[["Total Amount"]], big.mark   = ",")` Euros)** was observed in `r max_tbl$DISTRICT`, in `r lubridate::month(max_tbl$date, label = TRUE, abbr =FALSE)`.

```{r by_district}

by_district <- dta %>% 
  group_by(DISTRICT) %>% 
  summarise(
    'Total mortgages' = sum(Mortgages),
    'Total Value' = sum(`Total Amount`),
    'Value Per mortage' = sum(`Total Amount`)/ sum(Mortgages)
  ) %>% 
  arrange(desc(`Total Value`))


by_district %>% 
  kable(
    format.args = list(big.mark = ',')
    , caption = paste0("Total Value (€) by District - Year ", mortage_year)
  ) 


```

### A bar plot

```{r bar_plot}

by_district %>% 
  mutate(DISTRICT = fct_reorder(DISTRICT, `Total Value`)) %>% 
  ggplot(
    aes(DISTRICT, `Total Value`)
  ) +
  geom_col(width = 0.5, fill = "#4d6d82")+
  coord_flip()+
  #scale_y_continuous(labels=scales::dollar_format())+
  scale_y_continuous(labels=function(x) paste0(x/ 1000000, "M"))+
  theme_minimal()+
  labs(
  x = "", y = "",
  title = paste0("Total Mortage Value (€) by Distict - Year ", mortage_year)
  )

```

### A Line plot

```{r line_plot, fig.width=9}

dta %>% 
  ggplot(
    aes(date, `Total Amount`, colour = DISTRICT)
    )+
  geom_line()+
  scale_y_continuous(labels=function(x) paste0(x/ 1000000, "M"))+
  theme_minimal()+
  labs(
  x = "", y = "",
  title = paste0("Total Mortage Value (€) by Distict - Year ", mortage_year)
  )

```
